---
title: "MA617HW3_b-d"
output: 
  pdf_document:
    latex_engine: xelatex
date: "2024-09-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


a.
```{r}

# Load required libraries
library(data.table)
library(lubridate)

# Define the base URL components for NOAA data
file_root <- "https://www.ndbc.noaa.gov/view_text_file.php?filename=44013h"
tail <- ".txt.gz&dir=data/historical/stdmet/"

# Initialize an empty list to store data for each year
all_years_data <- list()

# Loop through each year from 1985 to 2023
for (year in 1985:2023) {
  
  # Construct the full URL for the current year
  path <- paste0(file_root, year, tail)
  
  # Try to process each year's data, and catch errors if they occur
  tryCatch({
    
    # Read the header line
    header <- scan(path, what = 'character', nlines = 1)
    
    # Determine how many lines to skip based on the year
    skip_lines <- if (year < 1990) 1 else 2
    
    # Read the buoy data, filling missing values and skipping the appropriate number of lines
    buoy_data <- fread(path, header = FALSE, skip = skip_lines, fill = TRUE)
    
    # Handle cases where the number of columns in data differs from the header
    if (length(header) != ncol(buoy_data)) {
      warning(paste("Mismatch in columns for year", year))
      
      # Adjust header or data if there is a mismatch
      if (length(header) > ncol(buoy_data)) {
        header <- header[1:ncol(buoy_data)]  # Trim the header
      } else {
        buoy_data <- buoy_data[, 1:length(header)]  # Trim the data
      }
    }
    
    # Assign column names to the data
    colnames(buoy_data) <- header
    
    # Validate the date columns (YY, MM, DD, hh) before parsing
    if (all(c("YY", "MM", "DD", "hh") %in% names(buoy_data))) {
      # Convert date components into a single Date column
      buoy_data$Date <- ymd_h(paste(buoy_data$YY, buoy_data$MM, buoy_data$DD, buoy_data$hh, sep = "-"))
    } else {
      warning(paste("Date columns missing for year", year))
      buoy_data$Date <- NA
    }
    
    # Store the buoy data for this year in the list
    all_years_data[[as.character(year)]] <- buoy_data
    
  }, error = function(e) {
    # If there's an error, print a message and continue with the next year
    warning(paste("Failed to process data for year", year, ":", e$message))
  })
}

# Combine all the data into a single data.table
combined_data <- rbindlist(all_years_data, use.names = TRUE, fill = TRUE)

# View the first few rows of the combined data
head(combined_data)

#save the csv file for 1985-2023
write.csv(combined_data, "buoy_44013_data_1985_2023.csv", row.names = FALSE)


nrow(combined_data)
```


b.
```{r}
# Load the CSV file
combined_data <- fread("buoy_44013_data_1985_2023.csv")

# List the columns where '999' or '99' represent missing values
missing_columns <- c("WDIR", "WSPD", "GST", "WVHT", "DPD", "APD", "MWD", "ATMP", "WTMP", "DEWP", "VIS")

# Replace '999' or '99' with NA in the relevant columns
combined_data[, (missing_columns) := lapply(.SD, function(x) fifelse(x == 999 | x == 99, NA, x)), .SDcols = missing_columns]

# Check if the replacement worked
summary(combined_data)

# Create a summary of NA counts for each column
na_counts <- sapply(combined_data, function(x) sum(is.na(x)))
print(na_counts)

# Check for NA values by Date (if you want to analyze them over time)
na_by_date <- combined_data[, .(NA_count = rowSums(is.na(.SD))), .SDcols = missing_columns, by = Date]

# Visualize the pattern of missing values over time using ggplot2
library(ggplot2)

ggplot(na_by_date, aes(x = Date, y = NA_count)) +
  geom_line() +
  labs(title = "Missing Values Over Time", x = "Date", y = "Number of Missing Values")




```
From the missing data graph, we see a high density of values unavailable. The range seems to fluctuate and has spikes showing there might have been external issues such as equipment malfunctioning and maintenance. However, from $1990 - 1995$ there seems to be less of the missing values, which means that the malfunctioning was less frequent.
From the summary that we see, WDIR, BAR, VIS, MWD seems to have a very high missing value, while VIS seems to have much lower missing values. This shows that some of the systems for specific criteria has been malfunctioning. 
Replacing missing values with NA might be effective to see the malfunctioning of systems, but it could be better for us to find other dataset that has specific criteria that we have more missing values on. There are some patterns shown in the graph, such as higher frequency of missing values in 1985 to 1990, which could be due to technological limitations. Also, there has been a significant increase in missing data from 1995 to 2023, showing possible operational changes or the end of dataset coverage. 

```{r}
# Load required libraries
library(data.table)
library(ggplot2)

# Filter out rows with missing values in key climate variables
climate_data <- combined_data[, .(Date, WTMP, ATMP, BAR, WSPD)]
climate_data <- climate_data[!is.na(WTMP) & !is.na(ATMP) & !is.na(BAR) & !is.na(WSPD)]

# Ensure the date column is properly formatted
climate_data$Date <- as.Date(climate_data$Date)

# Plot Water Temperature over Time
ggplot(climate_data, aes(x = Date, y = WTMP)) +
  geom_line(color = "blue") +
  geom_smooth(method = "lm", se = FALSE, color = "red") + # Trend line
  labs(title = "Water Temperature Over Time", x = "Year", y = "Water Temperature (°C)") +
  theme_minimal()

# Plot Atmospheric Temperature over Time
ggplot(climate_data, aes(x = Date, y = ATMP)) +
  geom_line(color = "orange") +
  geom_smooth(method = "lm", se = FALSE, color = "red") + # Trend line
  labs(title = "Atmospheric Temperature Over Time", x = "Year", y = "Atmospheric Temperature (°C)") +
  theme_minimal()

# Plot Barometric Pressure over Time with zoomed-in scale
ggplot(climate_data, aes(x = Date, y = BAR)) +
  geom_line(color = "green") +
  geom_smooth(method = "lm", se = FALSE, color = "red") +  # Trend line
  labs(title = "Barometric Pressure Over Time", x = "Year", y = "Barometric Pressure (hPa)") +
  scale_y_continuous(limits = c(950, 1050)) +  # Set realistic limits for barometric pressure
  theme_minimal()

# Plot Wind Speed over Time with adjusted scale
ggplot(climate_data, aes(x = Date, y = WSPD)) +
  geom_line(color = "purple") +
  geom_smooth(method = "lm", se = FALSE, color = "red") +  # Trend line
  labs(title = "Wind Speed Over Time", x = "Year", y = "Wind Speed (m/s)") +
  scale_y_continuous(limits = c(0, 25)) +  # Set realistic limits for wind speed
  theme_minimal()

# Convert Date to a numeric format (number of days since origin)
climate_data$Date_numeric <- as.numeric(climate_data$Date)

# Fit linear models for each variable
lm_wind_speed <- lm(WSPD ~ Date_numeric, data = climate_data)
lm_wtmp <- lm(WTMP ~ Date_numeric, data = climate_data)
lm_atmp <- lm(ATMP ~ Date_numeric, data = climate_data)
lm_bar <- lm(BAR ~ Date_numeric, data = climate_data)

# Extract slopes (coefficients for Date_numeric)
slope_wind_speed <- coef(lm_wind_speed)["Date_numeric"]
slope_wtmp <- coef(lm_wtmp)["Date_numeric"]
slope_atmp <- coef(lm_atmp)["Date_numeric"]
slope_bar <- coef(lm_bar)["Date_numeric"]

# Print the slopes
cat("Slope of Atmospheric Temperature Trend:", slope_atmp, "\n")
cat("Slope of Water Temperature Trend:", slope_wtmp, "\n")
cat("Slope of Barometric Pressure Trend:", slope_bar, "\n")
cat("Slope of Wind Speed Trend:", slope_wind_speed, "\n")

```


To see the effects of climate change, I used four different variables which were Atmospheric temperature, water temperature, barometric pressure, and wind speed. These were variables that I thought would represent climate changes, and hence plotted a time series plot. I also drew a trend line for each graphs to see if there were any increase or decrease in total. 
As you can see from the slopes, the atmospheric trend as well as water temperature and wind speed tredn tends to increase by a small amount. However, what looks like a small amount of increase can actually be a very big increase in the climate in general. Hence, I tried to divide the data into two, first half and second half, and see the differences in the mean. 



```{r}
library(data.table)
library(ggplot2)

# Filter out rows with missing values in key climate variables
climate_data <- combined_data[, .(Date, WTMP, ATMP, BAR, WSPD)]
climate_data <- climate_data[!is.na(WTMP) & !is.na(ATMP) & !is.na(BAR) & !is.na(WSPD)]

# Ensure the date column is properly formatted
climate_data$Date <- as.Date(climate_data$Date)

# Split the data into two halves
mid_date <- climate_data$Date[round(nrow(climate_data) / 2)]

first_half <- climate_data[Date <= mid_date]
second_half <- climate_data[Date > mid_date]

# Calculate the mean for each variable in both halves
means_first_half <- first_half[, lapply(.SD, mean, na.rm = TRUE), .SDcols = c("WTMP", "ATMP", "BAR", "WSPD")]
means_second_half <- second_half[, lapply(.SD, mean, na.rm = TRUE), .SDcols = c("WTMP", "ATMP", "BAR", "WSPD")]

# Combine the results for comparison
mean_comparison <- data.table(
  Variable = c("Water Temperature (WTMP)", "Atmospheric Temperature (ATMP)", "Barometric Pressure (BAR)", "Wind Speed (WSPD)"),
  First_Half_Mean = as.numeric(means_first_half),
  Second_Half_Mean = as.numeric(means_second_half),
  Difference = as.numeric(means_second_half) - as.numeric(means_first_half)
)

# Display the comparison
print(mean_comparison)

```

As you can see from the mean, there has been some significant changes in the temperature. For both water and atmospheric temperature, there has been almost 1 degree difference. However, unlike the slope showing an increase, there has been a decrease in the temperatures. This means that the mean could have been affected by in what weather there were missing data. If there were more missing data during the summer for temperatures, the temperatures could be decreasing while the slope still tends to increase. 


d.
```{r}
# Load necessary libraries
library(dplyr)

# Load the rainfall dataset
rainfall_data <- read.csv("Rainfall.csv")

# Check the structure of the dataset to verify column names
str(rainfall_data)

# Ensure the DATE column exists and is in proper format
# If DATE is in a different format, adjust the format string accordingly
rainfall_data$DATE <- as.Date(rainfall_data$DATE, format = "%Y-%m-%d")

# Summarize rainfall data: Total, Average, and Maximum Rainfall
rainfall_summary <- rainfall_data %>%
  summarise(
    Total_Rainfall = sum(HPCP, na.rm = TRUE),   # Sum of rainfall
    Average_Rainfall = mean(HPCP, na.rm = TRUE), # Average rainfall
    Max_Rainfall = max(HPCP, na.rm = TRUE)      # Maximum rainfall
  )

# Print the rainfall summary
print(rainfall_summary)

```
From the data frame, the data on total rainfall, average rainfall, and max rainfall is shown. 

```{r}
# Load necessary libraries
library(dplyr)
library(ggplot2)

# Load the dataset (make sure to replace with your actual file path)
rainfall_data <- read.csv("Rainfall.csv")

# Convert DATE to proper datetime format (assuming format is YYYYMMDD HH:MM)
rainfall_data$DATE <- as.POSIXct(rainfall_data$DATE, format = "%Y%m%d %H:%M")

# Summarize the dataset to check for missing or incorrect values
summary(rainfall_data)

# Check for negative or unusually large values in the rainfall data (HPCP)
rainfall_data <- rainfall_data %>%
  filter(HPCP >= 0 & HPCP <= 500)  # Assuming rainfall should be between 0 and 500 mm

# Aggregate total rainfall by day, month, and year for analysis
rainfall_data_daily <- rainfall_data %>%
  group_by(Day = as.Date(DATE)) %>%
  summarise(Daily_Rainfall = sum(HPCP, na.rm = TRUE))

rainfall_data_monthly <- rainfall_data %>%
  group_by(Month = format(DATE, "%Y-%m")) %>%
  summarise(Monthly_Rainfall = sum(HPCP, na.rm = TRUE))

# Fix the Month format for ggplot (convert to date format with first day of the month)
rainfall_data_monthly$Month <- as.Date(paste0(rainfall_data_monthly$Month, "-01"), format = "%Y-%m-%d")

rainfall_data_yearly <- rainfall_data %>%
  group_by(Year = format(DATE, "%Y")) %>%
  summarise(Yearly_Rainfall = sum(HPCP, na.rm = TRUE))

# View the summarized data
print(head(rainfall_data_daily))
print(head(rainfall_data_monthly))
print(head(rainfall_data_yearly))

# Plot rainfall patterns over time
# Daily Rainfall
ggplot(rainfall_data_daily, aes(x = Day, y = Daily_Rainfall)) +
  geom_line(color = "blue") +
  labs(title = "Daily Rainfall in Boston", x = "Date", y = "Rainfall (mm)") +
  theme_minimal()

# Monthly Rainfall (now with proper Date format)
ggplot(rainfall_data_monthly, aes(x = Month, y = Monthly_Rainfall)) +
  geom_line(color = "green") +
  labs(title = "Monthly Rainfall in Boston", x = "Month", y = "Rainfall (mm)") +
  theme_minimal()

# Yearly Rainfall
ggplot(rainfall_data_yearly, aes(x = as.numeric(Year), y = Yearly_Rainfall)) +
  geom_line(color = "red") +
  labs(title = "Yearly Rainfall in Boston (1985-2013)", x = "Year", y = "Rainfall (mm)") +
  theme_minimal()

```
With the analysis above, I chose to find the daily, monthly, and yearly rainfall. This shows a vague understanding of what the rain looks like in Boston over the years. I have also created a chart to see the numeric values as well as the graph for the trend. As you can see, here has been increasing rainfall in boston, when looking at the graph. 


```{r}
library(data.table)

# Load rainfall data
rainfall_data <- fread("Rainfall.csv")

# Load weather data
weather_data <- fread("buoy_44013_data_1985_2023.csv")

# Convert the DATE column in rainfall data to proper date format
rainfall_data$DATE <- as.Date(rainfall_data$DATE, format = "%Y%m%d")

# Ensure the Date column in weather data is in date format
weather_data$Date <- as.Date(weather_data$Date)

# Aggregating buoy data by Date
daily_weather <- weather_data[, .(
  Avg_Barometric_Pressure = mean(BAR, na.rm = TRUE),
  Avg_Wind_Speed = mean(WSPD, na.rm = TRUE),
  Avg_Atmospheric_Temperature = mean(ATMP, na.rm = TRUE)
), by = .(Date)]

# Merge the rainfall data with the weather data by the DATE column
merged_data <- merge(rainfall_data, daily_weather, by.x = "DATE", by.y = "Date", all.x = TRUE)

# Check the merged data
head(merged_data)

# Build a linear model to predict rainfall based on weather conditions
model <- lm(HPCP ~ Avg_Barometric_Pressure + Avg_Wind_Speed + Avg_Atmospheric_Temperature, data = merged_data)

# Summary of the model
summary(model)

# Make predictions (if needed)
predictions <- predict(model, newdata = merged_data)

# Evaluate model performance
library(Metrics)
mae_value <- mae(merged_data$HPCP, predictions)
rmse_value <- rmse(merged_data$HPCP, predictions)

cat("Mean Absolute Error (MAE):", mae_value, "\n")
cat("Root Mean Square Error (RMSE):", rmse_value, "\n")


```





```{r}
# Ensure weather_data and rainfall_data are correctly loaded

# Merge the two datasets by Date if the date columns match in both datasets
weather_data$Date <- as.Date(weather_data$Date)  # Convert to Date if necessary
rainfall_data$DATE <- as.Date(rainfall_data$DATE, format = "%Y-%m-%d")

# Merge the data
merged_data <- merge(weather_data, rainfall_data, by.x = "Date", by.y = "DATE")

# Simple regression model using weather variables to predict Rainfall
model <- lm(HPCP ~ BAR + WSPD + ATMP, data = merged_data)

# Summary of the model
summary(model)

# Predict rainfall using the model
merged_data$predicted_rainfall <- predict(model, newdata = merged_data)

# Plot actual vs predicted rainfall
library(ggplot2)

ggplot(merged_data, aes(x = predicted_rainfall, y = HPCP)) +
  geom_point(color = "blue", alpha = 0.5) +
  geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed") +  # Line y=x
  labs(title = "Actual vs Predicted Rainfall", x = "Predicted Rainfall (mm)", y = "Actual Rainfall (mm)") +
  theme_minimal()


```
For this model, I used pressure, wind speed, and atmospheric temperature to give out an assumption on how much rain will fall. Even though it does not seem to show a statistically significant outcome, it was a journey for me. 
I used the data from both of the csv files, using buoy data as my input variables, and rainfall as my output variable. I assumed that there will be some correlation between the rainfall and the changes in pressure, wind speed, and temperature. However, since weather is not easily predictable, the r^2 value or the actual and predicted rainfall did not show a significant result. I splitted the data to use some as predicted and as actual data, which shows the low accuracy of the model. However, as someone who feels like there should be more interesting areas in weather forecasting, this helped me to think about what could actually make people laugh. Giving a unprecise model, and guessing together with the people watching the weather forcast could be like a lottery!

